Analysis-ready VCF at Biobank scale using Zarr

### 2025-01-22 paper.tex
Add figure for network throughput

Finally, to explore the use of cloud computing services with VCF Zarr,
we uploaded the Zarr directory hierarchy for chromosome 2 to an AWS S3 
standard bucket. We repeated the benchmarks above the same subset
562,640 variants, and the same 
single-threaded sequential Zarr code.
Extracting the position array required 15s CPU, 23.8s elapsed.
The af-dist calculation required 2.70min CPU, 5.23min elapsed. 
Creating the filter mask
required 3.05min CPU, 11.15min elapsed.

While these benchmarks are comparable with results obtained from the 
shared file-system in the Genomics England research environment,
the sequential single-threaded access patterns used do not map
well to modern cloud computing environments.
As each request to a cloud object store has a high latency, it 
is important to have many requests running concurrently to achieve 
good performance~\citep{durner2023exploiting}.
To explore the limits of performance we developed a simple prototype
Zarr implementation, designed to maximise concurrency using asynchrous
IO and multiple processes (see Methods for details).
Fig~\ref{fig-s3-throughput} shows the rate at which genotypes 
can be decoded to RAM as well as rate at which we can compute
af-dist (using Numba to classify genotypes, as before)
as we increase the number of parallel processes on
an EC2 c5.24xlarge instance with 96 vCPUs (48 physical).
The maximum decompression rate of 47.4GiB/s 
occurs at 64 processes, with a corresponding 
network data rate of 
% FIXME MiB/s
127 MB/s, far below the capbility of the instance,
indicating that this process was compute-constrained.

@@

>>>>

Finally, to explore the use of cloud computing services with VCF Zarr,
we uploaded the Zarr directory hierarchy for chromosome 2 to an AWS S3 
standard bucket. We repeated the benchmarks above on the same subset
of 562,640 variants, using the same 
single-threaded sequential Zarr code.
Extracting the position array required 15s CPU, 23.8s elapsed.
The af-dist calculation required 2.70min CPU, 5.23min elapsed. 
Creating the filter mask
required 3.05min CPU, 11.15min elapsed.

While these benchmarks are comparable with results obtained on the 
shared file-system in the Genomics England research environment,
the sequential single-threaded access patterns used do not map
well to modern cloud computing environments.
As each request to a cloud object store has a high latency, it 
is important to have many requests running concurrently to achieve 
good performance~\citep{durner2023exploiting}.
To explore the limits of performance we developed a simple prototype
Zarr implementation, designed to maximise concurrency using asynchrous
IO and multiple processes (see Methods for details).
Fig~\ref{fig-s3-throughput} shows the rate at which genotypes 
can be decoded to RAM as well as rate at which we can compute
af-dist (using Numba to classify genotypes, as before)
as we increase the number of parallel processes on
an EC2 c5.24xlarge instance with 96 vCPUs (48 physical).
The maximum decompression rate of 47.4GiB/s 
occurs at 64 processes, with a corresponding 
network data rate of 127 MB/s [FIXME confirm and convert to MiB/s].
Fig~\ref{fig-s3-network-throughput} explores the limitations imposed by
network access by measuring the rate at which we can download the chunks
into RAM (without decompressing) using the same parallel and asynchronous
architecture. Here, the maximum throughput of

To assess the limitations imposed by network access, we repeated this
experiment but measured only the rate at which chunks can be downloaded

@@

\begin{figure}[h]
\includegraphics{figures/s3-network-throughput}
\caption{Chunk data download rate on AWS.
\label{fig-s3-network-throughput}}
\end{figure}
